---
title: "Unifying OCC across Time Periods"
author: "Clinton"
date: "12 April 2020"
output: html_document
editor_options: 
  chunk_output_type: console
---

<br>

```
Objective: To come up with a unified OCC code, allowing for comparisons of occupation counts across the three time periods of 1850, 1880 and 1910.

occ_label and occ1950 both had low number of levels (approximately 230) each, but were found to have levels nested within each other. I.e. there were 147 levels of occ_label in 1 level of occ1950. Some contradictions were also found.

Two approaches are proposed. (i) Clustering - to manually regroup occstr, (ii) NLP - to use tokenisation to define a dictionary of occupations which would be then used to label occstr
```

<br>

### 1. Set Up {.tabset .tabset-fade}

#### README

In general, the data sets are relatively large, and take up a large amount of memory after loading. It is recommended that after loading, that the data sets be subsetted or summarised, and unused data frames to be removed.

The code is shown in the next three tabs to show any changes that may have been made. In general, no columns or rows are removed. Filtering occured in the next step.

<br>

#### Loading 1850

```{r message = FALSE, warning = FALSE}
library(dplyr)
library(ggplot2)
library(treemap)
library(extrafont)
library(stringr)
loadfonts()
library(knitr)
library(kableExtra)
library(tidyr)
library(DT)
library(tidytext)

# Loading
mn_1850 = readr::read_csv("../Data/census_1850_occ_mn.csv") %>% mutate(city = "Manhattan")
bk_1850 = readr::read_csv("../Data/census_1850_occ_bk.csv") %>% mutate(city = "Brooklyn")
OCC_1880 = readr::read_csv("../Data/OCC1880.csv")
OCC_1950 = readr::read_csv("../Data/OCC1950.csv")

# Combining Data
combined_1850 = rbind(mn_1850, bk_1850) %>% 
  select(age, sex, marst, race, labforce,
         occ, city, occ1950, occstr) %>%
  left_join(OCC_1880, by = c("occ" = "code")) %>%
  left_join(OCC_1950, by = c("occ1950" = "code")) %>%
  mutate(race = factor(race,
                       levels = c(100,120,200,210,300),
                       labels = c("White", "Black (white)",
                                  "Black / African American",
                                  "Mulatto", "American Indian")),
         labforce = factor(labforce,
                           levels = c(0,1,2),
                           labels = c("N/A",
                                      "No, not in the labor force",
                                      "Yes, in the labor force")),
         sex = factor(sex,
                      levels = c(1,2),
                      labels = c("Male", "Female")),
         marst = factor(marst,
                        levels = c(1:6),
                        labels = c("Married, spouse present",
                                   "Married, spouse absent",
                                   "Separated",
                                   "Divorced",
                                   "Widowed",
                                   "Never married/single"))) %>%
  filter(labforce == "Yes, in the labor force")

# Theme setting for visualisations
theme_set(theme_minimal() + 
            theme(panel.grid.minor = element_blank(),
                  text = element_text(family = "Quicksand",
                                      colour = "black")))

# Source occ_modifier
source("0_OCC_Modifier.R")

# Saving Memory
rm(bk_1850)
rm(mn_1850)
```

<br>

#### Loading 1880

```{r message = FALSE, warning = FALSE}
# Loading
mn_1880 = readr::read_csv("../Data/census_1880_occ_mn.csv") %>% mutate(city = "Manhattan")
bk_1880 = readr::read_csv("../Data/census_1880_occ_bk.csv") %>% mutate(city = "Brooklyn")

# Combining Data
combined_1880 = rbind(mn_1880, bk_1880) %>% 
  select(age, sex, race, labforce,
         occ, city, occstr, occ1950) %>%
  left_join(OCC_1880, by = c("occ" = "code")) %>%
  mutate(age = ifelse(age == "Less than 1 year old",
                      0,
                      age),
         age = as.numeric(age)) %>%
  filter(labforce == "Yes, in the labor force")

rm(OCC_1880)
rm(bk_1880)
rm(mn_1880)
```

<br>

#### Loading 1910

```{r message = FALSE, warning = FALSE}
# Loading
mn_1910 = readr::read_csv("../Data/census_1910_occ_mn.csv") %>% mutate(city = "Manhattan")
bk_1910 = readr::read_csv("../Data/census_1910_occ_bk.csv") %>% mutate(city = "Brooklyn")
OCC_1950 = readr::read_csv("../Data/OCC1950.csv")

# Combining Data
combined_1910 = rbind(mn_1910, bk_1910) %>%
  left_join(OCC_1950, by = c("occ1950" = "code")) %>%
  select(age, sex, labor_force, race, occ1950, occstr, occ1950_label) %>%
  mutate(sex = factor(sex,
                      levels = c(1,2),
                      labels = c("Male", "Female")),
         labor_force = factor(labor_force,
                              levels = c(0,1,2),
                              labels = c("N/A", 
                                         "No, not in the labor force",
                                         "Yes, in the labor force"))) %>%
  filter(labor_force == "Yes, in the labor force")

rm(OCC_1950)
rm(mn_1910)
rm(bk_1910)
```

<br>

### 2. Visualisation of Inconsistencies

| Year | Variable | Count | Remarks | Information in metadata | 
| ----- | ----- | ----- | ----- | ----- |
| 1850 | occ | 262 | Code | Occupation |
| 1850 | occ1950 | 192 | Code | Occupation, 1950 basis |
| 1850 | occstr | 6641 | Actual occupation, not code | Occupation, alphabetic string |
| 1880 | occ | 274 | Code | Occupation |
| 1880 | occ1950 | 239 | Actual occupation, not code | Occupation, 1950 basis |
| 1880 | occstr | 79860 | Actual occupation, not code | Occupation, alphabetic string | 
| 1880 | occhisco | 628 | Actual occupation, not code | Occupation, HISCO classification |
| 1910 | occ1950 | 220 | Code | Occupation, 1950 basis |
| 1910 | occstr | 68725 | Actual occupation, not code | Occupation, alphabetic string |

From the table above, several options of using a consistent occupation coding scheme across the three years are available:
* Using `occ1950` which is consistent across all three years
* Using `occstr` which appears in all three years, but unsure if consistent across all three years
* Creating a new coding schema across all three years

<br>

### 3. Comparing `occ1950` to `occ` (`occ_label`) {.tabset .tabset-fade}

#### Possible Approaches

<br>

Given that `occ` has 268 unique occupations, while `occ1950` has 230 unique occupations, it suggests that several occupations in `occ` are grouped together into `occ1950` in 1880.

<br>

```{r}
combined_1880 %>% 
  filter(labforce == "Yes, in the labor force") %>% 
  select(occ1950, occ_label, occstr) %>%
  group_by(occ1950, occ_label) %>%
  summarise(count = n()) %>%
  group_by(occ1950) %>%
  mutate(occ_count = n()) %>%
  filter(occ_count > 1) %>%
  datatable()
```
<br>

Some observations are made:

* Some `occ` / `occ_label`s are nested within `occ1950`. The most common one is "Operative and kindred workers (n.e.c.)", which has 147 unique occ_labels such as "Wheelwrights" and "Wire makers and workers" nested within it.

* The inverse is also weirdly true, where the `occ_label` "Wood turners, carvers, and woodenware makers" has 3 `occ1950` nested within it - "Operative and kindred workers (n.e.c.)", "Service workers, except private household (n.e.c.)" and "Sawyers".

* Some obvious contradictions are also noted, such as in observations where the `occ1950` is "Stationary firemen", while the `occ_label` is "Sailors". It is noted that in this particular case the `occstr` was "MARINE FIREMAN"

* Finally, there appears to be some blanks in `occ_label`, but not in `occ1950`.

<br>

```{r}
combined_1880 %>% 
  group_by(occ1950, occ_label, occstr) %>%
  summarise(count = n()) %>%
  mutate(count_perc = round(count / NROW(combined_1880) * 100, 
                            digits = 2)) %>%
  arrange(desc(count)) %>%
  head(100) %>%            # Only 100 combinations are shown for ease of loading
  datatable()
```

* In coming up with an approach to find the "best" variable to encapsulate occupation, two approaches are proposed.

* The first approach would be akin to clustering, where we start working to group together `occstr`. For example, the `occstr` "LABORER IN STABLE" would be clustered into the largest "LABORER". Another example would be to group "KEEPING HOUSE" and "KEEPS HOUSE" together. This approach might work best by work on "higher value" levels first, where working to clean the top 100 unique combinations of `occ_label`, `occstr` and `occ_label` would cover about 53.9% of the entire data set. However, this would mean that efforts would taper out eventually, as eventually we end up manually clustering unique observations of only one count repeatedly.

* The second approach would be to use a natural language processing `bag-of-words` approach. We would come up with a new column for occupations, where they would be based on the most common words found within `occstr`. All words in `occstr` would be tokenised, and counted to obtain a summary of what the most common words are (i.e. "LABORER IN BREWERY", "LABORER IN COTTON STORE" have laborer as the most common words). Subsequently, set number of words can be extracted to use as a dictionary to label `occstr`. A list of the most common uni-grams are shown in the table below.

<br>

#### 1850

<br>

```{r}
combined_1850 %>%
  unnest_tokens(input = occstr,
                output = word,
                token = "words") %>%
  anti_join(stop_words) %>%
  group_by(word) %>%
  summarise(count = n()) %>%
  arrange(desc(count)) %>% 
  head(200) %>%             # Only 200 words are shown for ease of loading
  datatable()
```

```{r message = FALSE, warning = FALSE}
combined_1850 %>%
  group_by(occstr) %>%
  summarise(count = n()) %>%
  unnest_tokens(input = occstr,
                output = ngrams,
                token = "ngrams",
                n = 5,
                n_min = 1) %>%
  filter(!is.na(ngrams)) %>%
  arrange(desc(count)) %>% 
  head(200) %>%             # Only 200 words are shown for ease of loading
  datatable()
```

<br>

#### 1880

<br>

```{r}
combined_1880 %>%
  unnest_tokens(input = occstr,
                output = word,
                token = "words") %>%
  anti_join(stop_words) %>%
  group_by(word) %>%
  summarise(count = n()) %>%
  arrange(desc(count)) %>% 
  head(200) %>%             # Only 200 words are shown for ease of loading
  datatable()
```

<br>

```{r message = FALSE, warning = FALSE}
combined_1880 %>%
  group_by(occstr) %>%
  summarise(count = n()) %>%
  unnest_tokens(input = occstr,
                output = ngrams,
                token = "ngrams",
                n = 5,
                n_min = 1) %>%
  filter(!is.na(ngrams)) %>%
  arrange(desc(count)) %>% 
  head(200) %>%             # Only 200 words are shown for ease of loading
  datatable()
```

<br>

#### 1910

<br>

```{r}
combined_1910 %>%
  unnest_tokens(input = occstr,
                output = word,
                token = "words") %>%
  anti_join(stop_words) %>%
  group_by(word) %>%
  summarise(count = n()) %>%
  arrange(desc(count)) %>% 
  head(200) %>%             # Only 200 words are shown for ease of loading
  datatable()
```

<br>

```{r message = FALSE, warning = FALSE}
combined_1910 %>%
  group_by(occstr) %>%
  summarise(count = n()) %>%
  unnest_tokens(input = occstr,
                output = ngrams,
                token = "ngrams",
                n = 5,
                n_min = 1) %>%
  filter(!is.na(ngrams)) %>%
  arrange(desc(count)) %>% 
  head(200) %>%             # Only 200 words are shown for ease of loading
  datatable()
```

<br>

#### Combining across all three time periods

```{r}
combined_summary = combined_1850 %>%
  group_by(occstr) %>%
  summarise(count = n()) %>%
  rbind(combined_1880 %>%
          group_by(occstr) %>%
          summarise(count = n())) %>%
  rbind(combined_1910 %>%
          group_by(occstr) %>%
          summarise(count = n())) %>%
  group_by(occstr) %>%
  summarise(count = sum(count)) %>%
  unnest_tokens(input = occstr,
                output = ngrams,
                token = "ngrams",
                n = 5,
                n_min = 1) %>%
  filter(!is.na(ngrams)) %>%
  arrange(desc(count))

combined_summary %>% 
  head(200) %>%
  datatable()
```

```{r eval = FALSE}
readr::write_csv(combined_summary, path = "../Data/combined_summary.csv")
```

<br>

Inspecting the output from above, several observations are noted:

* There exists a large number of "makers" (i.e. dress makers), "clerk" "teacher", "driver", "merchants", "servant"
* There exists a number of mis-spellings, or inconsistent spellings, such as "laborer", "labor" and "labourer"

<br>

### 4. Application of `occ_modifier`

From `combined_summary`, `occ_modifier` was developed using the data frame as a reference. The function's main purpose is to identify `occstr`s that are nested within other `occstr`s (i.e. CLERK from CLERK IN STORE). It is noted that the identification and reclassification of `occstr`s give priority to `occstr`s that have a higher count. In other words, should there exist an `occstr` that contains two unique `occstr`s, it would be reclassified into the `occstr` with a higher count.

<br>

```{r}
pre1 = combined_1850 %>%
  group_by(occstr) %>%
  summarise(count = n()) %>%
  rbind(combined_1880 %>%
          group_by(occstr) %>%
          summarise(count = n())) %>%
  rbind(combined_1910 %>%
          group_by(occstr) %>%
          summarise(count = n())) %>%
  group_by(occstr) %>%
  summarise(count = sum(count)) %>%
  arrange(desc(count)) %>%
  head(200)

pre1$count %>% sum()
pre1 %>% datatable()
```

<br>

```{r}
pre2 = combined_1850 %>%
  group_by(occstr) %>%
  summarise(count = n()) %>%
  rbind(combined_1880 %>%
          group_by(occstr) %>%
          summarise(count = n())) %>%
  rbind(combined_1910 %>%
          group_by(occstr) %>%
          summarise(count = n())) %>%
  mutate(occstr = occ_modifier(occstr)) %>%
  group_by(occstr) %>%
  summarise(count = sum(count)) %>%
  arrange(desc(count)) %>%
  head(200)

pre2$count %>% sum()
pre2 %>% datatable()
```

<br>

Inspecting the top 200 occupations, it can be seen that the count of the top 200 occupations has increased from 1928171 to 2276802. Furthermore, some interesting changes would be that CLERK is now the occupation with the largest count, over taking laborer. 

<br>

### 5. Improvements {.tabset .tabset-fade}

#### 1850

```{r}
combined_1850 %>%
  group_by(occ_label, occ1950_label, occstr) %>%
  summarise(count = n()) %>%
  left_join(combined_1850 %>% 
              mutate(occstr = occ_modifier(occstr)) %>%
              group_by(occ_label, occ1950_label, occstr) %>%
              summarise(count_modified = n()),
            by = c("occ_label", "occ1950_label", "occstr")
            ) %>%
  mutate(improvement = paste0(round((count_modified - count) / 
                                      count * 100, digits = 1), 
                              "%")) %>%
  arrange(desc(count_modified)) %>%
  head(200) %>%
  datatable()
```

<br>

#### 1880

```{r}
combined_1880 %>%
  group_by(occ_label, occ1950, occstr) %>%
  summarise(count = n()) %>%
  left_join(combined_1880 %>% 
              mutate(occstr = occ_modifier(occstr)) %>%
              group_by(occ_label, occ1950, occstr) %>%
              summarise(count_modified = n()),
            by = c("occ_label", "occ1950", "occstr")
            ) %>%
  mutate(improvement = paste0(round((count_modified - count) / 
                                      count * 100, digits = 1), 
                              "%")) %>%
  arrange(desc(count_modified)) %>%
  head(200) %>%
  datatable()
```

<br>

#### 1910

```{r}
combined_1910 %>%
  group_by(occ1950_label, occstr) %>%
  summarise(count = n()) %>%
  left_join(combined_1910 %>% 
              mutate(occstr = occ_modifier(occstr)) %>%
              group_by(occ1950_label, occstr) %>%
              summarise(count_modified = n()),
            by = c("occ1950_label", "occstr")
            ) %>%
  mutate(improvement = paste0(round((count_modified - count) / 
                                      count * 100, digits = 1), 
                              "%")) %>%
  arrange(desc(count_modified)) %>%
  head(200) %>%
  datatable()
```

<br>