---
title: 'Standardizing Occupations Over Time'
author: "Dan Miller + Clinton Wong"
date: "4/28/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

```{r message = FALSE, warning = FALSE}
library(dplyr)
library(ggplot2)
library(treemap)
library(extrafont)
library(stringr)
loadfonts()
library(knitr)
library(kableExtra)
library(tidyr)
library(DT)
#### Loading 1850
# Loading
mn_1850 = readr::read_csv("../../Data/census_1850_occ_mn.csv") %>% mutate(city = "Manhattan")
bk_1850 = readr::read_csv("../../Data/census_1850_occ_bk.csv") %>% mutate(city = "Brooklyn")
OCC_1880 = readr::read_csv("../../Data/OCC1880.csv")

# Combining Data
combined_1850 = rbind(mn_1850, bk_1850) %>% 
  select(age, sex, marst, race, labforce,
         occ, city, everything()) %>%
  left_join(OCC_1880, by = c("occ" = "code")) %>%
  mutate(race = factor(race,
                       levels = c(100,120,200,210,300),
                       labels = c("White", "Black (white)",
                                  "Black / African American",
                                  "Mulatto", "American Indian")),
         labforce = factor(labforce,
                           levels = c(0,1,2),
                           labels = c("N/A",
                                      "No, not in the labor force",
                                      "Yes, in the labor force")),
         sex = factor(sex,
                      levels = c(1,2),
                      labels = c("Male", "Female")),
         marst = factor(marst,
                        levels = c(1:6),
                        labels = c("Married, spouse present",
                                   "Married, spouse absent",
                                   "Separated",
                                   "Divorced",
                                   "Widowed",
                                   "Never married/single")))

# Theme setting for visualisations
theme_set(theme_minimal() + 
            theme(panel.grid.minor = element_blank(),
                  text = element_text(family = "Arial",
                                      colour = "black")))
col_sex = c("#e21737", "#0b648f")
col_miss = c("#86b817", "#e53238")

# Saving Memory
rm(bk_1850)
rm(mn_1850)
```

```{r message = FALSE, warning = FALSE}
#### Loading 1880
# Loading
mn_1880 = readr::read_csv("../../Data/census_1880_occ_mn.csv") %>% mutate(city = "Manhattan")
bk_1880 = readr::read_csv("../../Data/census_1880_occ_bk.csv") %>% mutate(city = "Brooklyn")
OCC_1880 = readr::read_csv("../../Data/OCC1880.csv")

# Combining Data
combined_1880 = rbind(mn_1880, bk_1880) %>% 
  select(age, sex, race, labforce,
         occ, city, everything()) %>%
  left_join(OCC_1880, by = c("occ" = "code")) %>%
  mutate(age = ifelse(age == "Less than 1 year old",
                      0,
                      age),
         age = as.numeric(age))

rm(bk_1880)
rm(mn_1880)
```

```{r message = FALSE, warning = FALSE}
#### Loading 1910
# Loading
mn_1910 = readr::read_csv("../../Data/census_1910_occ_mn.csv") %>% mutate(city = "Manhattan")
bk_1910 = readr::read_csv("../../Data/census_1910_occ_bk.csv") %>% mutate(city = "Brooklyn")
OCC_1950 = readr::read_csv("../../Data/OCC1950.csv")
OCC_1920 = readr::read_csv("../../Data/OCC1920.csv")
IND_1950 = readr::read_csv("../../Data/IND1950.csv")

# Combining Data
combined_1910 = rbind(mn_1910, bk_1910) %>%
  left_join(OCC_1950, by = c("occ1950" = "code")) %>%
  mutate(sex = factor(sex,
                      levels = c(1,2),
                      labels = c("Male", "Female")),
         labor_force = factor(labor_force,
                              levels = c(0,1,2),
                              labels = c("N/A", 
                                         "No, not in the labor force",
                                         "Yes, in the labor force")))

rm(mn_1910)
rm(bk_1910)
```
## Overview Tables

#### Exploring the relationship between coded variables: `occ`, `occ1950` and occupational strings: `occstr` over time.

| Year | Variable | Count | Remarks | Information in metadata | 
| ----- | ----- | ----- | ----- | ----- |
| 1850 | occ | 262 | Code | Occupation |
| 1850 | occ1950 | 192 | Code | Occupation, 1950 basis |
| 1850 | occstr | 6641 | Actual occupation, not code | Occupation, alphabetic string |
| 1880 | occ | 274 | Code | Occupation |
| 1880 | occ1950 | 239 | Actual occupation, not code | Occupation, 1950 basis |
| 1880 | occstr | 79860 | Actual occupation, not code | Occupation, alphabetic string | 
| 1880 | occhisco | 628 | Actual occupation, not code | Occupation, HISCO classification |
| 1910 | occ1950 | 220 | Code | Occupation, 1950 basis |
| 1910 | occstr | 68725 | Actual occupation, not code | Occupation, alphabetic string |

### 1. Unique Strings by Primary Occupational Codes {.tabset .tabset-fade}

#### README

```{r message = FALSE, warning = FALSE}
#Generating tables of unique OCCSTR per OCC/OCC1950 for each census year
str_count_1880 <- combined_1880 %>% 
    group_by(label) %>% 
    count(occstr) %>% 
    left_join(OCC_1880) %>% 
    rename(occ = label)

str_count_1850 <- combined_1850 %>% 
    group_by(label) %>% 
    count(occstr) %>% 
    left_join(OCC_1880) %>% 
    rename(occ = label)

str_count_1910 <- combined_1910 %>% 
    group_by(label) %>% 
    count(occstr) %>% 
    left_join(OCC_1950) %>% 
    rename(occ1950 = label)
```

#### 1850
```{r message = FALSE, warning = FALSE}
str_count_1850 %>% 
    datatable(options = list(pageLength = 1000, scrollY = "400px"))
```

#### 1880
```{r message = FALSE, warning = FALSE}
str_count_1880 %>% 
    datatable(options = list(pageLength = 1000, scrollY = "400px"))
```

#### 1910
```{r message = FALSE, warning = FALSE}

str_count_1910 %>% 
    datatable(options = list(pageLength = 1000, scrollY = "400px"))

```

### 2. Comparing OCC and OCC1950 Over Time {.tabset .tabset-fade}

#### README

#### 1850
```{r}
combined_1850 %>% 
  filter(labforce == "Yes, in the labor force") %>% 
  select(occ1950, label, occstr) %>%
  left_join(OCC_1950, by = c("occ1950" = "code")) %>% 
  select(-occ1950) %>% 
  rename(occ = label.x, occ1950 = label.y) %>% 
  group_by(occ1950, occ) %>%
  summarise(count = n()) %>%
  group_by(occ1950) %>%
  mutate(occ_count = n()) %>%
  filter(occ_count > 1) %>%
  datatable(options = list(pageLength = 1000, scrollY = "400px"))
```

#### 1880
```{r}
combined_1880 %>% 
  filter(labforce == "Yes, in the labor force") %>% 
  select(occ1950, label, occstr) %>%
  rename(occ = label) %>% 
  group_by(occ1950, occ) %>%
  summarise(count = n()) %>%
  group_by(occ1950) %>%
  mutate(occ_count = n()) %>%
  filter(occ_count > 1) %>%
  datatable(options = list(pageLength = 1000, scrollY = "400px"))
```

#### 1910

This comparison isn't possible with 1910 census data, lacking `OCC` though we may eventually be able to compare `OCC1920` with the `OCC1950` coding scheme. 


### 3. Combinations of OCC and OCC1950 with related OCCSTR {.tabset .tabset-fade}

#### README

#### 1850
```{r}
combined_1850 %>% 
    filter(labforce == "Yes, in the labor force") %>% 
    select(occ1950, label, occstr) %>% 
    left_join(OCC_1950, by = c("occ1950" = "code")) %>% 
    select(-occ1950) %>% 
    rename(occ = label.x, occ1950 = label.y) %>% 
    group_by(occ1950, occ) %>%
    mutate(occstr_merge = paste0(unique(occstr), collapse = ", ")) %>% 
    group_by(occ1950, occ, occstr_merge) %>% 
    summarise(count = n()) %>%
    group_by(occ1950) %>%
    mutate(occ_count = n()) %>%
    filter(occ_count > 1) %>% 
    datatable(options = list(pageLength = 1000, scrollY = "400px"))
```

#### 1880
```{r message = FALSE, warning = FALSE}
combined_1880 %>% 
    filter(labforce == "Yes, in the labor force") %>% 
    select(occ1950, label, occstr) %>% 
    rename(occ = label) %>% 
    group_by(occ1950, occ) %>%
    mutate(occstr_merge = paste0(unique(occstr), collapse = ", ")) %>% 
    group_by(occ1950, occ, occstr_merge) %>% 
    summarise(count = n()) %>%
    group_by(occ1950) %>%
    mutate(occ_count = n()) %>%
    filter(occ_count > 1) %>% 
    datatable(options = list(pageLength = 1000, scrollY = "400px"))
```

#### 1910
```{r message = FALSE, warning = FALSE}
combined_1910 %>% 
    filter(labor_force == "Yes, in the labor force") %>% 
    group_by(occ1950) %>%
    mutate(occstr_merge = paste0(unique(occstr), collapse = ", ")) %>% 
    count(occstr_merge) %>% 
    left_join(OCC_1950, by = c("occ1950" = "code")) %>% 
    ungroup() %>% 
    select(-occ1950) %>% 
    rename(occ1950 = label) %>% 
    select(occ1950, occstr_merge, n) %>% 
    datatable(options = list(pageLength = 1000, scrollY = "400px"))
```


### 4. Unique OCCSTR with related combinations of OCC and OCC1950 {.tabset .tabset-fade}

#### README

#### 1850
```{r message = FALSE, warning = FALSE}
combined_1850 %>% 
     filter(labforce == "Yes, in the labor force") %>% 
     select(occ1950, label, occstr) %>% 
     left_join(OCC_1950, by = c("occ1950" = "code")) %>% 
     select(-occ1950) %>% 
     rename(occ = label.x, occ1950 = label.y) %>% 
     group_by(occstr) %>% 
     mutate(occ_merge = paste0(unique(occ), collapse = "; ")) %>% 
     mutate(occ1950_merge = paste0(unique(occ1950), collapse = "; ")) %>%
     mutate(n_occ = n_distinct(occ)) %>% 
     mutate(n_occ1950 = n_distinct(occ1950)) %>% 
     group_by(occ_merge, occ1950_merge, occstr, n_occ1950, n_occ) %>% 
     summarise(count = n()) %>% 
     datatable(options = list(pageLength = 1000, scrollY = "400px"))
```

#### 1880
```{r message = FALSE, warning = FALSE}
combined_1880 %>% 
     filter(labforce == "Yes, in the labor force") %>% 
     select(occ1950, label, occstr) %>% 
     rename(occ = label) %>% 
     group_by(occstr) %>% 
     mutate(occ_merge = paste0(unique(occ), collapse = "; ")) %>% 
     mutate(occ1950_merge = paste0(unique(occ1950), collapse = "; ")) %>%
     mutate(n_occ = n_distinct(occ)) %>% 
     mutate(n_occ1950 = n_distinct(occ1950)) %>% 
     group_by(occ_merge, occ1950_merge, occstr, n_occ1950, n_occ) %>% 
     summarise(count = n()) %>% 
     datatable(options = list(pageLength = 1000, scrollY = "400px"))
```

#### 1910
```{r message = FALSE, warning = FALSE}
combined_1910 %>% 
    filter(labor_force == "Yes, in the labor force") %>%
    select(label, occstr) %>% 
    rename(occ1950 = label) %>% 
    group_by(occstr) %>% 
    mutate(occ1950_merge = paste0(unique(occ1950), collapse = "; ")) %>%
    mutate(n_occ1950 = n_distinct(occ1950)) %>% 
    group_by(occ1950_merge, occstr, n_occ1950) %>% 
    summarise(count = n()) %>% 
    datatable(options = list(pageLength = 1000, scrollY = "400px"))
```

## Industry Codes + Merged Tables

### 1. OCC, OCC1950, Industry with related OCCSTR  {.tabset .tabset-fade}

#### README

#### 1850
```{r message = FALSE, warning = FALSE}
combined_1850 %>% 
    filter(labforce == "Yes, in the labor force") %>% 
    select(occ1950, label, occstr, ind1950) %>% 
    left_join(OCC_1950, by = c("occ1950" = "code")) %>% 
    select(-occ1950) %>%
    left_join(IND_1950, by = c("ind1950" = "Code")) %>% 
    select(-ind1950) %>% 
    rename(occ = label.x, occ1950 = label.y, ind1950 = Label) %>% 
    group_by(occ1950, occ, ind1950) %>%
    mutate(occstr_merge = paste0(unique(occstr), collapse = ", "), occstr_count = n_distinct(occstr)) %>% 
    group_by(occ1950, occ, ind1950, occstr_merge, occstr_count) %>% 
    summarise(n = n()) %>%
    group_by(occ1950, occ) %>%
    mutate(ind_count = n()) %>%
    group_by(occ1950) %>% 
    mutate(occ_count = n()) %>% 
    select(occ1950, occ, ind1950, occstr_merge, occstr_count, occ_count, ind_count, n) %>%     
    datatable(options = list(pageLength = 1000, scrollY = "400px"))
```


#### 1880
```{r message = FALSE, warning = FALSE}
combined_1880 %>% 
    filter(labforce == "Yes, in the labor force") %>% 
    select(occ1950, label, ind1950, occstr) %>% 
    rename(occ = label) %>% 
    group_by(occ1950, occ, ind1950) %>%
    mutate(occstr_merge = paste0(unique(occstr), collapse = ", "), occstr_count = n_distinct(occstr)) %>% 
    group_by(occ1950, occ,ind1950, occstr_merge, occstr_count) %>% 
    summarise(n = n()) %>%
    group_by(occ1950, occ) %>%
    mutate(ind_count = n()) %>% 
    group_by(occ1950) %>% 
    mutate(occ_count = n()) %>% 
    select(occ1950, occ, ind1950, occstr_merge, occstr_count, occ_count, ind_count, n) %>% 
    datatable(options = list(pageLength = 1000, scrollY = "400px"))
```


#### 1910
```{r message = FALSE, warning = FALSE}

combined_1910 %>% 
    filter(labor_force == "Yes, in the labor force") %>% 
    group_by(occ1950, ind_1950) %>%
    mutate(occstr_merge = paste0(unique(occstr), collapse = ", ")) %>% 
    count(occstr_merge) %>% 
    left_join(OCC_1950, by = c("occ1950" = "code")) %>% 
    left_join(IND_1950, by = c("ind_1950" = "Code")) %>% 
    ungroup() %>% 
    select(-occ1950, -ind_1950) %>% 
    rename(occ1950 = label, ind1950 = Label) %>% 
    select(occ1950, ind1950, occstr_merge, n) %>% 
    datatable(options = list(pageLength = 1000, scrollY = "400px"))
```

### 2. Unique OCCSTR with related combinations of OCC and OCC1950 {.tabset .tabset-fade}

#### README

#### 1850
```{r message = FALSE, warning = FALSE}
combined_1850 %>% 
    filter(labforce == "Yes, in the labor force") %>% 
    select(occ1950, label, occstr, ind1950) %>% 
    left_join(OCC_1950, by = c("occ1950" = "code")) %>% 
    select(-occ1950) %>%
    left_join(IND_1950, by = c("ind1950" = "Code")) %>% 
    select(-ind1950) %>% 
    rename(occ = label.x, occ1950 = label.y, ind1950 = Label) %>% 
     group_by(occstr) %>% 
     mutate(occ_merge = paste0(unique(occ), collapse = "; "), occ_count = n_distinct(occ)) %>% 
     mutate(occ1950_merge = paste0(unique(occ1950), collapse = "; "), occ1950_count = n_distinct(occ1950)) %>%
    mutate(ind_merge = paste0(unique(ind1950), collapse = "; "), ind_count = n_distinct(ind1950)) %>% 
     group_by(occstr, occ_merge, occ_count, occ1950_merge, occ1950_count, ind_merge, ind_count) %>% 
     summarise(n = n()) %>% 
     datatable(options = list(pageLength = 1000, scrollY = "400px"))
```

#### 1880
```{r message = FALSE, warning = FALSE}
combined_1880 %>% 
     filter(labforce == "Yes, in the labor force") %>% 
     select(occ1950, label, ind1950, occstr) %>% 
     rename(occ = label) %>% 
     group_by(occstr) %>% 
     mutate(occ_merge = paste0(unique(occ), collapse = "; "), occ_count = n_distinct(occ)) %>% 
     mutate(occ1950_merge = paste0(unique(occ1950), collapse = "; "), occ1950_count = n_distinct(occ1950)) %>%
    mutate(ind_merge = paste0(unique(ind1950), collapse = "; "), ind_count = n_distinct(ind1950)) %>% 
     group_by(occstr, occ_merge, occ_count, occ1950_merge, occ1950_count, ind_merge, ind_count) %>% 
     summarise(count = n()) %>% 
     datatable(options = list(pageLength = 1000, scrollY = "400px"))
```

#### 1910
```{r message = FALSE, warning = FALSE}
combined_1910 %>% 
    filter(labor_force == "Yes, in the labor force") %>%
    select(label, ind_1950, occstr) %>% 
    left_join(IND_1950, by = c("ind_1950" = "Code")) %>%
    select(-ind_1950) %>% 
    rename(occ1950 = label, ind1950 = Label) %>% 
    group_by(occstr) %>% 
    mutate(occ1950_merge = paste0(unique(occ1950), collapse = "; "), occ1950_count = n_distinct(occ1950)) %>%
    mutate(ind_merge = paste0(unique(ind1950), collapse = "; "), ind_count = n_distinct(ind1950)) %>% 
    group_by(occstr, occ1950_merge, occ1950_count, ind_merge, ind_count) %>% 
    summarise(count = n()) %>% 
    datatable(options = list(pageLength = 1000, scrollY = "400px"))
```


<!-- ### 3. OCCSTR over time (merged) with related OCC, OCC1950, and INDUSTRY -->

## Recoding Problems

### 1. Exploring Blank or Not Yet Classified records {.tabset .tabset-fade}

#### README

#### 1850


```{r message = FALSE, warning = FALSE}
combined_1850 %>% 
    filter(labforce == "Yes, in the labor force") %>% 
    select(occ1950, label, occstr) %>% 
    left_join(OCC_1950, by = c("occ1950" = "code")) %>% 
    select(-occ1950) %>% 
    rename(occ = label.x, occ1950 = label.y) %>% 
    filter(occ == "Blank") %>% 
    group_by(occ1950, occstr) %>% 
    summarise(count = n()) %>% 
    datatable(options = list(pageLength = 1000, scrollY = "400px"))
```

#### 1880
```{r message = FALSE, warning = FALSE}
combined_1880 %>% 
    filter(labforce == "Yes, in the labor force") %>% 
    select(occ1950, label, occstr) %>% 
    rename(occ = label) %>% 
    filter(occ == "Blank") %>% 
    group_by(occ1950, occstr, occ) %>%
    summarise(count = n()) %>% 
    datatable(options = list(pageLength = 1000, scrollY = "400px"))
```

#### 1910
```{r message = FALSE, warning = FALSE}
combined_1910 %>% 
    filter(labor_force == "Yes, in the labor force") %>% 
    select(label, occstr) %>% 
    rename(occ1950 = label) %>% 
    filter(occ1950 == "Not yet classified") %>% 
    group_by(occ1950, occstr) %>% 
    summarise(count = n()) %>% 
    datatable(options = list(pageLength = 1000, scrollY = "400px"))
```

#### 1910 (w/ Industry)

### 2. Testing Extension of OCC via OCCSTR Crosswalk {.tabset .tabset-fade}

#### README

#### MISSING

Based on OCCSTR join
Next steps would include running this again after initial OCCSTR improvements and clarifying schema for conflicting / multiple code-string matched (would industry help?)
```{r message = FALSE, warning = FALSE}
join_1910 <-  combined_1910 %>% 
    select(occ1950, occstr, label) %>% 
    group_by(occ1950, occstr, label) %>% 
    summarise(count = n())
```

```{r message = FALSE, warning = FALSE}
##create occ to join 1880 code by 1880 string

occ_join <- combined_1880 %>% 
     filter(labforce == "Yes, in the labor force") %>% 
     select(label, occstr) %>% 
     rename(occ = label) %>% 
     group_by(occstr) %>% 
     mutate(occ_merge = paste0(unique(occ), collapse = "; "), occ_count = n_distinct(occ)) %>% 
     group_by(occstr, occ_merge, occ_count) %>% 
     summarise(count = n())
```

```{r message = FALSE, warning = FALSE}
#Loading joins on OCCSTR for recoding OCC
#matched
#1880 first (eventually merge all of the joined tables)
#shows the occ1950 categories that the strings are coded into
#many multiple matches, pointing to the signifiance of intersecting these joins with other fields (industry)
matched_1910_occ1880 <- readr::read_csv("../../Data/matching_1910/Census1910_matched-occ-codes-from-1880.csv")

#missing
#1880 first (eventually merge all of the joined tables)
missing_1910_occ1880 <- readr::read_csv("../../Data/matching_1910/Census1910_missing-occ-codes-in-1880.csv") %>% 
    left_join(join_1910, by = c("Occupation (string)" = "occstr"))
```

```{r message = FALSE, warning = FALSE}
missing_1910_occ1880 %>% 
    select(occ1950 = label,count, occstr = "Occupation (string)", N) %>% 
    datatable(options = list(pageLength = 1000, scrollY = "400px"))
```

#### MATCHING

Based on OCCSTR join
Next steps would include running this again after initial OCCSTR improvements and clarifying schema for conflicting / multiple code-string matched (would industry help?)
```{r message = FALSE, warning = FALSE}
matched_1910_occ1880 %>% 
    select(occstr, N) %>% 
    filter(occstr != "Missing") %>% 
    datatable(options = list(pageLength = 1000, scrollY = "400px"))
```

#### MATCHING with new OCC (1880)

Based on OCCSTR join

```{r message = FALSE, warning = FALSE}
matched_1910_occ1880 %>% 
    select(occstr) %>% 
    filter(occstr != "Missing") %>%
    left_join(occ_join, by = c("occstr" = "occstr")) %>% 
    datatable(options = list(pageLength = 1000, scrollY = "400px"))
```



<!-- #### MISSING (w/ Industry) -->
<!-- ```{r message = FALSE, warning = FALSE} -->

<!-- ``` -->

<!-- #### MATCHING (w/ Industry) -->
<!-- ```{r message = FALSE, warning = FALSE} -->

<!-- ``` -->




