---
title: "HNYC - Census Subcommittee - Occupation Counts"
author: ""
date: "4/10/2020"
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r echo=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

```{r message = FALSE, warning = FALSE}
library(dplyr)
library(ggplot2)
library(treemap)
library(extrafont)
library(stringr)
loadfonts()
library(knitr)
library(kableExtra)
library(tidyr)

#### Loading 1850

# Loading
mn_1850 = readr::read_csv("../../Data/census_1850_occ_mn.csv") %>% mutate(city = "Manhattan")
bk_1850 = readr::read_csv("../../Data/census_1850_occ_bk.csv") %>% mutate(city = "Brooklyn")
OCC_1880 = readr::read_csv("../../Data/OCC1880.csv")

# Combining Data
combined_1850 = rbind(mn_1850, bk_1850) %>% 
  select(age, sex, marst, race, labforce,
         occ, city, everything()) %>%
  left_join(OCC_1880, by = c("occ" = "code")) %>%
  mutate(race = factor(race,
                       levels = c(100,120,200,210,300),
                       labels = c("White", "Black (white)",
                                  "Black / African American",
                                  "Mulatto", "American Indian")),
         labforce = factor(labforce,
                           levels = c(0,1,2),
                           labels = c("N/A",
                                      "No, not in the labor force",
                                      "Yes, in the labor force")),
         sex = factor(sex,
                      levels = c(1,2),
                      labels = c("Male", "Female")),
         marst = factor(marst,
                        levels = c(1:6),
                        labels = c("Married, spouse present",
                                   "Married, spouse absent",
                                   "Separated",
                                   "Divorced",
                                   "Widowed",
                                   "Never married/single")))

# Theme setting for visualisations
theme_set(theme_minimal() + 
            theme(panel.grid.minor = element_blank(),
                  text = element_text(family = "Arial",
                                      colour = "black")))
col_sex = c("#e21737", "#0b648f")
col_miss = c("#86b817", "#e53238")

# Saving Memory
rm(OCC_1880)
rm(bk_1850)
rm(mn_1850)
```

```{r message = FALSE, warning = FALSE}
#### Loading 1880
# Loading
mn_1880 = readr::read_csv("../../Data/census_1880_occ_mn.csv") %>% mutate(city = "Manhattan")
bk_1880 = readr::read_csv("../../Data/census_1880_occ_bk.csv") %>% mutate(city = "Brooklyn")
OCC_1880 = readr::read_csv("../../Data/OCC1880.csv")

# Combining Data
combined_1880 = rbind(mn_1880, bk_1880) %>% 
  select(age, sex, race, labforce,
         occ, city, everything()) %>%
  left_join(OCC_1880, by = c("occ" = "code")) %>%
  mutate(age = ifelse(age == "Less than 1 year old",
                      0,
                      age),
         age = as.numeric(age))

rm(OCC_1880)
rm(bk_1880)
rm(mn_1880)
```

```{r message = FALSE, warning = FALSE}
#### Loading 1910
# Loading
mn_1910 = readr::read_csv("../../Data/census_1910_occ_mn.csv") %>% mutate(city = "Manhattan")
bk_1910 = readr::read_csv("../../Data/census_1910_occ_bk.csv") %>% mutate(city = "Brooklyn")
OCC_1950 = readr::read_csv("../../Data/OCC1950.csv")

# Combining Data
combined_1910 = rbind(mn_1910, bk_1910) %>%
  left_join(OCC_1950, by = c("occ1950" = "code")) %>%
  mutate(sex = factor(sex,
                      levels = c(1,2),
                      labels = c("Male", "Female")),
         labor_force = factor(labor_force,
                              levels = c(0,1,2),
                              labels = c("N/A", 
                                         "No, not in the labor force",
                                         "Yes, in the labor force")))

rm(OCC_1950)
rm(mn_1910)
rm(bk_1910)
```

### Occupation Counts and Percentages for each HNYC Census Year {.tabset .tabset-fade}

#### README

The tabs in this section are organized by Census year and contain corresponding charts that show observed counts and percentages per occupational code.For these charts, occupations displayed for 1850 and 1880 are coded by IPUMS to an 1880 basis, while the 1910 complete count Census dataset is coded by IPUMS to a 1950 basis.

Note that the graphs are all produced using `occ_label` (according to IPUMS coding scheme for that census year) as the key occupational variable, not `occstr` or the occupation string as enumerated/transcribed.

#### 1850 {.tabset .tabset-fade}

##### Graph of All Occupations by Count

```{r fig.height = 50, fig.width = 10}
combined_1850 %>%
  filter(labforce == "Yes, in the labor force") %>%
  count(label) %>%
  mutate(perc_lf = round(n / sum(combined_1850$labforce == "Yes, in the labor force") * 100,
                         digits = 2),
         text = paste0(n," (", perc_lf, "%)")) %>%
  
  ggplot(aes(y = n, x = reorder(label, n))) +
  geom_col(fill = "#f9a541") + 
  coord_flip() + 
  geom_text(aes(label = text),
            hjust = 0,
            family = "Arial") + 
  scale_y_continuous(breaks = seq(0, 25000, 5000),
                     expand = expand_scale(mult = c(0,0.2))) +
  
  labs(x = "Occupation", y = "Count",
       title = "Graph of All Occupations by Count in 1850") +
  theme(panel.grid.major.y = element_blank(),
        legend.text = element_text(colour = "black"))
```

<br>

##### Graph of All Occupations by Count (Breaks)

Red dashed lines indicate natural breaks (10) in the distrubtion of observed and coded occupations.

```{r fig.height = 50, fig.width = 10}
plot1 <- combined_1850 %>%
  filter(labforce == "Yes, in the labor force") %>%
  count(label) %>%
  mutate(perc_lf = round(n / sum(combined_1850$labforce == "Yes, in the labor force") * 100,
                         digits = 2),
         text = paste0(n," (", perc_lf, "%)")) %>%
 
  ggplot(aes(y = n, x = reorder(label, n))) +
  geom_col(fill = "#f9a541") + 
  coord_flip() + 
  geom_text(aes(label = text),
            hjust = 0,
            family = "Arial") + 
  scale_y_continuous(breaks = seq(0, 25000, 5000),
                     expand = expand_scale(mult = c(0,0.2))) +
  
  labs(x = "Occupation", y = "Count",
       title = "Graph of All Occupations by Count in 1850") +
  theme(panel.grid.major.y = element_blank(),
        legend.text = element_text(colour = "black"))

occ_1850_normalized <- combined_1850 %>%
  filter(labforce == "Yes, in the labor force") %>%
  count(label) %>%
  mutate(perc_lf = round(n / sum(combined_1850$labforce == "Yes, in the labor force") * 100,
                         digits = 2),
         text = paste0(n," (", perc_lf, "%)"))

#classInt::classIntervals(occ_1850_normalized$perc_lf, 6)
#classInt::classIntervals(occ_1850_normalized$perc_lf, n = 10, style = "jenks")
#classInt::classIntervals(occ_1850_normalized$n, n = 10, style = "jenks")
#classInt::classIntervals(occ_1850_normalized$n, 5)

#quantile(occ_1850_normalized$perc_lf)

plot1 + 
      geom_vline(aes(xintercept = 142.5), color = "red",
            linetype = "dashed", size = 0.5)+ 
      geom_vline(aes(xintercept = 182.5), color = "red",
            linetype = "dashed", size = 0.5) + 
      geom_vline(aes(xintercept = 217.5), color = "red",
            linetype = "dashed", size = 0.5)+ 
      geom_vline(aes(xintercept = 232.5), color = "red",
            linetype = "dashed", size = 0.5)+ 
      geom_vline(aes(xintercept = 240.5), color = "red",
            linetype = "dashed", size = 0.5)+ 
      geom_vline(aes(xintercept = 246.5), color = "red",
            linetype = "dashed", size = 0.5)+ 
      geom_vline(aes(xintercept = 249.5), color = "red",
            linetype = "dashed", size = 0.5)+ 
      geom_vline(aes(xintercept = 252.5), color = "red",
            linetype = "dashed", size = 0.5)+ 
      geom_vline(aes(xintercept = 253.5), color = "red",
            linetype = "dashed", size = 0.5)

```

<br>

##### Graph of All Occupations by Type

```{r fig.height = 50, fig.width = 10}
combined_1850 %>%
  filter(labforce == "Yes, in the labor force") %>%
  count(label) %>%
  mutate(perc_lf = round(n / sum(combined_1850$labforce == "Yes, in the labor force") * 100,
                         digits = 2),
         text = paste0(n," (", perc_lf, "%)")) %>%
  
  ggplot(aes(y = n, x = label, n)) +
  geom_col(fill = "#f9a541") + 
  coord_flip() + 
  geom_text(aes(label = text),
            hjust = 0,
            family = "Arial") + 
  scale_y_continuous(breaks = seq(0, 25000, 5000),
                     expand = expand_scale(mult = c(0,0.2))) +
  
  labs(x = "", y = "Count",
       title = "Graph of All Occupations by Type in 1850") +
  theme(panel.grid.major.y = element_blank(),
        legend.text = element_text(colour = "black"))
```

<br>

#### 1880 {.tabset .tabset-fade}


##### Graph of All Occupations by Count

```{r fig.height = 50, fig.width = 10}
combined_1880 %>%
  filter(labforce == "Yes, in the labor force" & 
           label != "Blank") %>%
  count(label) %>%
  mutate(perc_lf = round(n / sum(combined_1880$labforce == "Yes, in the labor force") * 100,
                         digits = 2),
         text = paste0(n," (", perc_lf, "%)")) %>%
  
  ggplot(aes(y = n, x = reorder(label, n))) +
  geom_col(fill = "#bc5090") + 
  coord_flip() + 
  geom_text(aes(label = text),
            hjust = 0,
            family = "Arial") + 
  scale_y_continuous(breaks = seq(0, 70000, 10000),
                     expand = expand_scale(mult = c(0,0.2))) +
  
  labs(x = "", y = "Count",
       title = "Graph of All Occupations by Count in 1880") +
  theme(panel.grid.major.y = element_blank(),
        legend.text = element_text(colour = "black"))
```

<br>

##### Graph of All Occupations by Count (Breaks)

Red dashed lines indicate natural breaks (10) in the distrubtion of observed and coded occupations.

```{r fig.height = 50, fig.width = 10}
plot2 <- combined_1880 %>%
  filter(labforce == "Yes, in the labor force" & 
           label != "Blank") %>%
  count(label) %>%
  mutate(perc_lf = round(n / sum(combined_1880$labforce == "Yes, in the labor force") * 100,
                         digits = 2),
         text = paste0(n," (", perc_lf, "%)")) %>%
  
  ggplot(aes(y = n, x = reorder(label, n))) +
  geom_col(fill = "#bc5090") + 
  coord_flip() + 
  geom_text(aes(label = text),
            hjust = 0,
            family = "Arial") + 
  scale_y_continuous(breaks = seq(0, 70000, 10000),
                     expand = expand_scale(mult = c(0,0.2))) +
  
  labs(x = "", y = "Count",
       title = "Graph of All Occupations by Count in 1880") +
  theme(panel.grid.major.y = element_blank(),
        legend.text = element_text(colour = "black"))

occ_1880_normalized <- combined_1880 %>%
  filter(labforce == "Yes, in the labor force" & 
           label != "Blank") %>%
  count(label) %>%
  mutate(perc_lf = round(n / sum(combined_1880$labforce == "Yes, in the labor force") * 100,
                         digits = 2),
         text = paste0(n," (", perc_lf, "%)"))

#classInt::classIntervals(occ_1880_normalized$perc_lf, n=10, style = "jenks")

plot2 +
     geom_vline(aes(xintercept = 161.5), color = "red",
            linetype = "dashed", size = 0.5) + 
      geom_vline(aes(xintercept = 215.5), color = "red",
            linetype = "dashed", size = 0.5) + 
      geom_vline(aes(xintercept = 238.5), color = "red",
            linetype = "dashed", size = 0.5) + 
      geom_vline(aes(xintercept = 250.5), color = "red",
            linetype = "dashed", size = 0.5) + 
      geom_vline(aes(xintercept = 259.5), color = "red",
            linetype = "dashed", size = 0.5) + 
      geom_vline(aes(xintercept = 261.5), color = "red",
            linetype = "dashed", size = 0.5) + 
      geom_vline(aes(xintercept = 263.5), color = "red",
            linetype = "dashed", size = 0.5) + 
      geom_vline(aes(xintercept = 264.5), color = "red",
            linetype = "dashed", size = 0.5) + 
      geom_vline(aes(xintercept = 265.5), color = "red",
            linetype = "dashed", size = 0.5)
    
```

<br>

##### Graph of All Occupations by Type

```{r fig.height = 50, fig.width = 10}
combined_1880 %>%
  filter(labforce == "Yes, in the labor force") %>%
  count(label) %>%
  mutate(perc_lf = round(n / sum(combined_1880$labforce == "Yes, in the labor force") * 100,
                         digits = 2),
         text = paste0(n," (", perc_lf, "%)")) %>%
  
  ggplot(aes(y = n, x = label, n)) +
  geom_col(fill = "#bc5090") + 
  coord_flip() + 
  geom_text(aes(label = text),
            hjust = 0,
            family = "Arial") + 
  scale_y_continuous(breaks = seq(0, 70000, 10000),
                     expand = expand_scale(mult = c(0,0.2))) +
  
  labs(x = "", y = "Count",
       title = "Graph of All Occupations by Type in 1880") +
  theme(panel.grid.major.y = element_blank(),
        legend.text = element_text(colour = "black"))
```

<br>

#### 1910 {.tabset .tabset-fade}

##### Graph of All Occupations by Count

```{r fig.height = 50, fig.width = 10}
combined_1910 %>%
  filter(labor_force == "Yes, in the labor force" & 
           label != "Not yet classified") %>%
  count(label) %>%
  mutate(perc_lf = round(n / sum(combined_1910$labor_force == "Yes, in the labor force") * 100,
                         digits = 2),
         text = paste0(n," (", perc_lf, "%)")) %>%
  
  ggplot(aes(y = n, x = reorder(label, n))) +
  geom_col(fill = "#003f5c") + 
  coord_flip() + 
  geom_text(aes(label = text),
            hjust = 0,
            family = "Arial") + 
  scale_y_continuous(breaks = seq(0, 200000, 20000),
                     expand = expand_scale(mult = c(0,0.25))) +
  
  labs(x = "", y = "Count",
       title = "Graph of All Occupations by Count in 1910") +
  theme(panel.grid.major.y = element_blank(),
        legend.text = element_text(colour = "black"))
```

<br>

##### Graph of All Occupations by Count (Breaks)

Red dashed lines indicate natural breaks (10) in the distrubtion of observed and coded occupations.

```{r fig.height = 50, fig.width = 10}
plot3 <- combined_1910 %>%
  filter(labor_force == "Yes, in the labor force" & 
           label != "Not yet classified") %>%
  count(label) %>%
  mutate(perc_lf = round(n / sum(combined_1910$labor_force == "Yes, in the labor force") * 100,
                         digits = 2),
         text = paste0(n," (", perc_lf, "%)")) %>%
  
  ggplot(aes(y = n, x = reorder(label, n))) +
  geom_col(fill = "#003f5c") + 
  coord_flip() + 
  geom_text(aes(label = text),
            hjust = 0,
            family = "Arial") + 
  scale_y_continuous(breaks = seq(0, 200000, 20000),
                     expand = expand_scale(mult = c(0,0.25))) +
  
  labs(x = "", y = "Count",
       title = "Graph of All Occupations by Count in 1910") +
  theme(panel.grid.major.y = element_blank(),
        legend.text = element_text(colour = "black"))

occ_1910_normalized <- combined_1910 %>%
  filter(labor_force == "Yes, in the labor force" & 
           label != "Not yet classified") %>%
  count(label) %>%
  mutate(perc_lf = round(n / sum(combined_1910$labor_force == "Yes, in the labor force") * 100,
                         digits = 2),
         text = paste0(n," (", perc_lf, "%)"))

#classInt::classIntervals(occ_1910_normalized$perc_lf, 10, style = "jenks")

plot3 +
    geom_vline(aes(xintercept = 121.5), color = "red",
            linetype = "dashed", size = 0.5) + 
      geom_vline(aes(xintercept = 162.5), color = "red",
            linetype = "dashed", size = 0.5) + 
      geom_vline(aes(xintercept = 185.5), color = "red",
            linetype = "dashed", size = 0.5) + 
      geom_vline(aes(xintercept = 194.5), color = "red",
            linetype = "dashed", size = 0.5) + 
      geom_vline(aes(xintercept = 199.5), color = "red",
            linetype = "dashed", size = 0.5) + 
      geom_vline(aes(xintercept = 203.5), color = "red",
            linetype = "dashed", size = 0.5) + 
      geom_vline(aes(xintercept = 205.5), color = "red",
            linetype = "dashed", size = 0.5) + 
      geom_vline(aes(xintercept = 206.5), color = "red",
            linetype = "dashed", size = 0.5) + 
      geom_vline(aes(xintercept = 209.5), color = "red",
            linetype = "dashed", size = 0.5)

```


<br>

##### Graph of All Occupations by Type

```{r fig.height = 50, fig.width = 10}
combined_1910 %>%
  filter(labor_force == "Yes, in the labor force" & 
           label != "Not yet classified") %>%
  count(label) %>%
  mutate(perc_lf = round(n / sum(combined_1910$labor_force == "Yes, in the labor force") * 100,
                         digits = 2),
         text = paste0(n," (", perc_lf, "%)")) %>%
  
  ggplot(aes(y = n, x = label, n)) +
  geom_col(fill = "#003f5c") + 
  coord_flip() + 
  geom_text(aes(label = text),
            hjust = 0,
            family = "Arial") + 
  scale_y_continuous(breaks = seq(0, 200000, 20000),
                     expand = expand_scale(mult = c(0,0.25))) +
  
  labs(x = "", y = "Count",
       title = "Graph of All Occupations by Type in 1910") +
  theme(panel.grid.major.y = element_blank(),
        legend.text = element_text(colour = "black"))
```

<br>
